Practical Hilbert space approximate Bayesian Gaussian processes for probabilistic programming

### 2020-04-24 Paper/manuscript_v2/manuscript_07_02.tex
remove unnecessary figures and commented text. Update file name

\usepackage{listings}
%\usepackage{slashbox}
\usepackage{caption}
\captionsetup{font=footnotesize}
\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black]{hyperref}

%\bibliographystyle{authordate1}
% TODO macro
\newcommand{\todo}[1]{\textcolor{blue}{\textbf{[#1]}}}
\newcommand{\resp}[1]{\textcolor{red}{\textbf{[#1]}}}

@@

%\tableofcontents

@@

Sparse GPs are based on low-rank approximations of the covariance matrix. The low-rank approximation with $m \ll n$ {\it inducing points} implies reduced memory requirements of $O(nm)$ and corresponding computational complexity of $O(nm^2)$. A unifying view on sparse GPs based on approximate generative methods is provided by \citet{quinonero2005unifying}, while a general review is provided by \citet{rasmussen2006gaussian}. \citet{Burt+Rasmussen+vanderWilk:2019} show that for regression with normally distributed covariates in $D$ dimensions and using the squared exponential covariance function, $M=O(\log^DN)$ is sufficient for an accurate approximation.
% More recent developments in the context of sparse GPs include a structured kernel interpolation method \citep{wilson2015kernel}. More recently, \citep{bui2017unifying} proposed to use inducing points-based sparse approximation methods to perform GP approximations at inference time rather than at modeling time.
An alternative class of low-rank approximations is based on forming a basis function approximation with $m \ll n$ basis functions. The basis functions are usually presented explicitly, but can also be used to form a low-rank covariance matrix approximation. Common basis function approximations rest on the spectral analysis and series expansions of GPs \citep{loeve1977probability,trees1968detection,adler1981geometry,cramer2013stationary}. Sparse spectrum GPs are based on a sparse approximation to the frequency domain representation of a GP \citep{lazaro2010sparse,quia2010sparse,gal2015improving}. Recently, \citet{hensman2017variational} presented a variational Fourier feature approximation for GPs that was derived for the Mat{\'e}rn class of kernels. Another related method for approximating kernels relies on random Fourier features  \citep{rahimi2008random,rahimi2009weighted}. Certain spline smoothing basis functions are equivalent to GPs with certain covariance functions \citep{wahba1990spline,Furrer+Nychka:2007}. Recent related work based on a spectral representation of GPs as an infinite series expansion with the Karhunen-Loève representation \citep[see, e.g.,][]{grenander1981abstract} is presented by \citet{JSSv090i10}.
% , but not all spline models are GPs (e.g., some spline models are not marginalization consistent).

%{\color{blue} Paul: Can we clarify how our discussed method relates to the former paragraph,
%which introduces all sorts of basis function approximations. I.e. is our method
%just one out of many or what makes it special? Perhaps we can somehow use the information provided
%in: While Sparse Spectrum GP is based on a sparse spectrum, the reduced-rank method proposed in this paper aims to make the spectrum as ‘full’ as possible at a given rank. Recent Splines models can reproduce the Matern family of covariance functions (see, e.g., \cite{wood2003thin}), however our approach can reproduce basically all of the stationary covariance functions.}

@@

%\textcolor{red}{Aki: I moved this from section 2, needs to be combined with other intro}
%We propose an approximate framework for fast and accurate inference
%for GPs. Using a basis function expansion, we
%approximate the GPs with a linear model. This
%representation has three main advantages: 1) it makes inference
%considerably faster due to the linear structure, 2) it is simple to
%implement, which makes it easy to use GPs as building
%blocks in more complicated models and can be used as latent function
%in non-Gaussian observational models allowing modelling flexibility,
%3) it can be made arbitrary accurate and the trade-off between
%computational complexity and approximation accuracy can easily be
%controlled.

@@

%In this paper, we analyze in detail the performance and accuracy of the method in relation to key factors such as the number of basis functions, desired prediction space, or properties of the true functional relationship between covariates and response variable. We provide intuitive visualizations and practical recommendations for the choice of these factors, which will help users to improve computational performance while maintaining close approximation to exact GPs.

@@

In this work, GPs will take the role of a prior distribution over function spaces for non-parametric latent functions in a Bayesian setting. Consider a data set $\mathcal{D} = \left\lbrace \bm{x}_n, y_n \right\rbrace_{n=1}^N$, where $y_n$ is modelled conditionally as $p(y_n \mid f(\bm{x}_n),\phi)$, where $p$ is some parametric distribution with parameters  $\phi$, and $f$ is an unknown function with GP prior, which depends on an input $\bm{x}_n\in {\rm I\!R}^D$. This generalizes readily to more complex models depending on several unknown functions, for example such as $p(y_n \mid f(\bm{x}_n),g(\bm{x}_n))$ or multilevel models. Our goal is to obtain the posterior distribution for the value of the function $\tilde{f}=f(\tilde{\bm{x}})$  evaluated at a new input point $\tilde{\bm{x}}$.
% That is, we want to obtain the predictive distribution $p(f^* \mid \mathcal{D})$ of $f^*$ conditioned on the data $\mathcal{D}$.\\

We assume a GP prior for $f \sim \GP(\mu(\bm{x}), k(\bm{x}, \bm{x}'))$, where $\mu: {\rm I\!R}^D \to {\rm I\!R}$ and $k: {\rm I\!R}^D \times {\rm I\!R}^D \to {\rm I\!R}$ are the mean and covariance functions, respectively,
%
\begin{align*}
 	\mu(\bm{x}) &= \mathbb{E}\!\left[f(\bm{x})\right],\\ 
 	k(\bm{x}, \bm{x}') &= \mathbb{E}\!\left[\left( f(\bm{x}) - \mu(\bm{x}) \right)\left( f(\bm{x}') - \mu(\bm{x}') \right)\right].
\end{align*} 
%
The mean and covariance functions completely characterize the GP prior, and control the a priori behavior of the function $f$. Let $\bm{f}=\left\lbrace f(\bm{x}_n) \right\rbrace_{n=1}^N$, then the resulting prior distribution for $\bm{f}$ is a multivariate Gaussian distribution $\bm{f} \sim \Normal(\bm{\mu}, \bm{K})$, where $\bm{\mu} = \left\lbrace \mu(\bm{x}_n) \right\rbrace_{n=1}^N$ is the mean and $\bm{K}$ the covariance matrix, where $K_{i,j}=k(\bm{x}_i,\bm{x}_j)$. In the following, we focus on zero-mean Gaussian processes, that is set $\mu(\bm{x}) = 0$. The covariance function $k(\bm{x}, \bm{x}')$ might depend on a set of hyperparameters, $\bm{\theta}$, but we will not write this dependency explicitly to ease the notation. The joint distribution of $\bm{f}$ and a new $\tilde{f}$ is also a multivariate Gaussian as,
%
\begin{align*}
p(\bm{f}, \tilde{f})=\Normal \left( \left[ \begin{array}{cc}
\bm{f} \\ 
f^*
\end{array} \right] \,\middle|\, \bm{0},\left[ \begin{array}{cc}
\bm{K}_{\bm{f},\bm{f}} & \bm{k}_{\bm{f},\tilde{f}} \\ 
\bm{k}_{\tilde{f},\bm{f}} & k_{\tilde{f},\tilde{f}}
\end{array} \right] \right),
\end{align*}
%
where $\bm{k}_{\bm{f},\tilde{f}}$ is the covariance between $\bm{f}$ and $\tilde{f}$, and $k_{\tilde{f},\tilde{f}}$ is the prior variance of $\tilde{f}$. %By using the conditioning properties of multivariate Gaussian distributions, we can derive the predictive distribution for $f^*$  given $\bm{f}$ analytically,
% \begin{align*}
	% p(f^*| \bm{f}) = \Normal(f^*| \bm{k}_{f^*,\bm{f}} \bm{K}_{\bm{f},\bm{f}}^{-1} \bm{f},  k_{f^*,f^*}-\bm{k}_{f^*,\bm{f}} \bm{K}_{\bm{f},\bm{f}}^{-1} \bm{k}_{\bm{f},f^*}  )
% \end{align*}

% {\color{red} Aki: the following until ``While Gaussian'' seems too basic to be in the paper:}

% The joint distribution of the observations $\bm{y} = \left\lbrace y_n \right\rbrace_{n=1}^N$ and function values $\bm{f}$ and $f^*$, $p(\bm{y}, \bm{f}, f^*)$, is the product of the conditional distribution for $\bm{y}$ given $\bm{f}$ and the joint distribution for $\bm{f}$ and $f^*$,
% \begin{align*}
% 	p(\bm{y}, \bm{f}, f^*) = p(\bm{y}|\bm{f})p(\bm{f}, f^*)
% \end{align*}

% By marginalizing over $\bm{f}$ and conditioning on vector of observations $\bm{y}$, we obtain the posterior distribution of interest
% %
% \begin{align}\label{eq:posterior}
% 	p(f^*|\bm{y}) = \frac{\int p(\bm{y}|\bm{f})p(\bm{f}, f^*)\mathrm{d}\bm{f}}{p(\bm{y})},
% \end{align}
% where $p(\bm{y})$ is the marginal likelihood and is given by
% %
% \begin{align}\label{eq:evidence}
% 	p(\textbf{y}) = \int p(\bm{y}|\bm{f})p(\bm{f}, f^*) \mathrm{d}\bm{f}\mathrm{d}f^*.
% \end{align}
% %
% If the observation model $p(\bm{y}|\bm{f})$ is Gaussian, both integrals in eq. \eqref{eq:posterior} and \eqref{eq:evidence} can be solved analytically conditioned on the hyperparameters. For example, an isotropic Gaussian likelihood yields the following closed-form solution
% \begin{align*}
% 	p(f^*| \bm{y}) = \Normal(f^*| \textbf{k}_{f^*,\bm{f}} (\textbf{K}_{\bm{f},\bm{f}} + \sigma^2 \bm{I})^{-1} \bm{f},  k_{f^*,f^*}-\bm{k}_{f^*,\bm{f}} (\bm{K}_{\bm{f},\bm{f}} + \sigma^2 \bm{I})^{-1} \bm{k}_{\bm{f},f^*}  ),
% \end{align*}
% where $\sigma^2$ is the noise variance.

@@

\begin{figure}[!t]
\centering
%\subfigure{\includegraphics[scale=0.38, trim = 10mm 100mm 5mm 30mm, clip]{ch5_fig1_legend_01.pdf}}\\
%\vspace{-9mm}
\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 10mm 0mm, clip]{ch5_fig1_Post_J.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 0mm 5mm 10mm 0mm, clip]{ch5_fig1_Sigma_J.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{ch5_fig1_Cov_J.pdf}}
%\subfigure{\includegraphics[scale=0.38, trim = 29mm 48mm 124mm 10mm, clip]{ch5_fig1_legend.pdf}}
\vspace{-1mm}
\caption{Mean posterior predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model (dashed red line) and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.}
  \label{fig1_Post_J}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 10mm 0mm, clip]{ch5_fig2_Post_L.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 0mm 5mm 10mm 0mm, clip]{ch5_fig2_Sigma_L.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{ch5_fig2_Cov_L.pdf}}
%\subfigure{\includegraphics[scale=0.38, trim = 29mm 48mm 124mm 10mm, clip]{ch5_fig2_legend.pdf}}
\vspace{-1mm}
\caption{Mean posterior predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model (dashed red line) and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.}
  \label{fig2_Post_L}
\end{figure}

@@

\begin{figure}
\centering
\begin{tabular}{ c c c }
\arrayrulecolor{darkgray}\hline
c = 1.05 &
\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part1.pdf}
\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part1.pdf} 
\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part1.pdf} & 
\multirow{32}{1cm}{ \includegraphics[scale=0.35, trim = 28mm 30mm 100mm 30mm, clip]{ch5_fig3_legend.pdf}}\\ 
\arrayrulecolor{lightgray}\cline{1-2}
c = 1.1 &
\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part2.pdf} 
\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part2.pdf} 
\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part2.pdf} &\\
\cline{1-2}
c = 1.2 &
\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part3.pdf} 
\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part3.pdf} 
\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part3.pdf} &\\
\cline{1-2}
c = 1.5 &
\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part4.pdf} 
\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part4.pdf} 
\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part4.pdf} & \\
\cline{1-2}
c = 2 &
\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part5.pdf} 
\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part5.pdf} 
\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part5.pdf} & \\
\cline{1-2}
c = 2.5 &
\includegraphics[scale=0.215, trim = 0mm 4mm 0mm 14mm, clip]{ch5_fig3_Post_part6.pdf}
\includegraphics[scale=0.215, trim = 0mm 4mm 0mm 14mm, clip]{ch5_fig3_Sigma_part6.pdf}
\includegraphics[scale=0.215, trim = 0mm 4mm 5mm 14mm, clip]{ch5_fig3_Cov_part6.pdf} &\\
\arrayrulecolor{darkgray}\hline
\end{tabular}
\caption{Posterior mean predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$. %The dashed red line of the exact GP can hardly be seen in the some of the plots because it is under the blue lines.
}
  \label{fig3_Post_part1}
\end{figure}

@@

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.42, trim = 10mm 100mm 5mm 30mm, clip]{ch5_fig7_legend.pdf}}\\
\vspace{-3mm}
\subfigure{\includegraphics[scale=0.31, trim = 1mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part1.pdf}}
\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part2.pdf}}
\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part3.pdf}}\\
\vspace{-3mm}
\subfigure{\includegraphics[scale=0.31, trim = 1mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part4.pdf}}
\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part5.pdf}}
\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part6.pdf}}\\
\vspace{-3mm}
\subfigure{\includegraphics[scale=0.31, trim = 1mm 5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part7.pdf}}
\subfigure{\includegraphics[scale=0.31, trim = 21mm 5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part8.pdf}}
\subfigure{\includegraphics[scale=0.31, trim = 21mm 5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part9.pdf}}\\
\caption{Mean posterior predictions of both exact GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic lengthscales ($\ell$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{true function}). %The dashed red line of the exact GP can hardly be seen in the plots because it is almost perfectly overlayed by the blue lines.
}
  \label{fig7_posterior_varing_lscale_part1}
\end{figure}

@@

%This fact reveals that choosing optimal values for the number of basis functions and boundary factor, looking at the recommendations and diagnosis provided by Figure \ref{fig6_lscale_vs_J_vs_c}, is essential to avoid a excessive computational time, especially in multivariate input spaces.

@@

The component $f_1(t)$ represents the long-term trends modeled by a GP with squared exponential covariance function,
%
\begin{equation*}
f_1(t) \sim \GP(0,k_1), \hspace{5mm} k_1(t,t') = \alpha_1 \exp\!\!\left(\!-\frac{1}{2} \frac{(t-t')^2}{\ell_1^2}\right), 
\end{equation*}
%
which means the function values $\bm{f}_1=\{f_1(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $\bm{K}_1$, where $K_{1_{t,s}}=k_1(t,s)$, with $t,s=1,\dots,T$. $\alpha_1$ and $\ell_1$ represent the marginal variance and lengthscale, respectively, of this GP prior component. 
The component $f_2(t)$ represents the yearly smooth seasonal pattern, using a periodic squared exponential covariance function (with period 365.25 to match the average length of the year) in a GP model,
%
\begin{equation*}
f_2(t) \sim \GP(0,k_2), \hspace{5mm} k_2(t,t') = \alpha_2 \exp\!\!\left(\!-\frac{2\,\sin^{\!2}\!(\pi(t-t')/365.25}{\ell_2^2}\right).
\end{equation*}
% which means the function values $\bm{f}_2=\{f_2(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $K_2$, where $K_{2_{t,s}}=k_2(t,s)$, with $t,s=1,\dots,T$.

The component $f_3(t)$ represents the weekly smooth pattern using a periodic squared exponential covariance function (with period 7 of length of the week) in a GP model,
%
\begin{equation*}
f_3(t) \sim \GP(0,k_3), \hspace{5mm} k_3(t,t') = \alpha_3 \exp\!\!\left(\!-\frac{2\,\sin^{\!2}\!(\pi(t-t')/7}{\ell_3^2}\right). 
\end{equation*}
% which means the function values $\bm{f}_3=\{f_3(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $K_3$, where $K_{3_{t,s}}=k_3(t,s)$, with $t,s=1,\dots,T$.

The component $f_4(t)$ represents the special days effects, modeled as a horse-shoe prior model \citep{carvalho2010,piironen2017sparsity}:
%
\begin{equation*}
f_4(t) \sim \Normal(0,\lambda^2_t \tau^2), \hspace{5mm} \lambda^2_t \sim \mathcal{C}^{+}(0,1).
\end{equation*}
%
A horse-shoe prior allows for sparse distributed effects. Its global parameter $\tau$ pulls all the weights (effects) globally towards zero, while the thick half-Cauchy tails for the local scales $\lambda_t$ allow some of the weights to escape the shrinkage. Different levels of sparsity can be accommodated by changing the value of $\tau$: for large $\tau$ all the variables have very diffuse prior distributions with very little shrinkage towards zero, but letting $\tau \rightarrow 0$ will shrink all the weights $f_4(t)$ to zero \citep{piironen2016hyperprior}. 
%More details about the specified priors for $\tau$ and $\lambda_t$ have been included in Appendix \ref{app_horse_shoe}.

@@

%In terms of scaling, one main drawback of the method is the scaling with respect to the number of input dimensions, as the number of basis functions $m$ scales exponentially with the number of dimensions.

>>>>

\usepackage{listings}
\usepackage{caption}
\captionsetup{font=footnotesize}
\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black,breaklinks=true]{hyperref}

@@

@@

Sparse GPs are based on low-rank approximations of the covariance matrix. The low-rank approximation with $m \ll n$ {\it inducing points} implies reduced memory requirements of $O(nm)$ and corresponding computational complexity of $O(nm^2)$. A unifying view on sparse GPs based on approximate generative methods is provided by \citet{quinonero2005unifying}, while a general review is provided by \citet{rasmussen2006gaussian}. \citet{Burt+Rasmussen+vanderWilk:2019} show that for regression with normally distributed covariates in $D$ dimensions and using the squared exponential covariance function, $M=O(\log^DN)$ is sufficient for an accurate approximation.
An alternative class of low-rank approximations is based on forming a basis function approximation with $m \ll n$ basis functions. The basis functions are usually presented explicitly, but can also be used to form a low-rank covariance matrix approximation. Common basis function approximations rest on the spectral analysis and series expansions of GPs \citep{loeve1977probability,trees1968detection,adler1981geometry,cramer2013stationary}. Sparse spectrum GPs are based on a sparse approximation to the frequency domain representation of a GP \citep{lazaro2010sparse,quia2010sparse,gal2015improving}. Recently, \citet{hensman2017variational} presented a variational Fourier feature approximation for GPs that was derived for the Mat{\'e}rn class of kernels. Another related method for approximating kernels relies on random Fourier features  \citep{rahimi2008random,rahimi2009weighted}. Certain spline smoothing basis functions are equivalent to GPs with certain covariance functions \citep{wahba1990spline,Furrer+Nychka:2007}. Recent related work based on a spectral representation of GPs as an infinite series expansion with the Karhunen-Loève representation \citep[see, e.g.,][]{grenander1981abstract} is presented by \citet{JSSv090i10}.

@@

@@

@@

In this work, GPs will take the role of a prior distribution over function spaces for non-parametric latent functions in a Bayesian setting. Consider a data set $\mathcal{D} = \left\lbrace \bm{x}_n, y_n \right\rbrace_{n=1}^N$, where $y_n$ is modelled conditionally as $p(y_n \mid f(\bm{x}_n),\phi)$, where $p$ is some parametric distribution with parameters  $\phi$, and $f$ is an unknown function with GP prior, which depends on an input $\bm{x}_n\in {\rm I\!R}^D$. This generalizes readily to more complex models depending on several unknown functions, for example such as $p(y_n \mid f(\bm{x}_n),g(\bm{x}_n))$ or multilevel models. Our goal is to obtain the posterior distribution for the value of the function $\tilde{f}=f(\tilde{\bm{x}})$  evaluated at a new input point $\tilde{\bm{x}}$.

We assume a GP prior for $f \sim \GP(\mu(\bm{x}), k(\bm{x}, \bm{x}'))$, where $\mu: {\rm I\!R}^D \to {\rm I\!R}$ and $k: {\rm I\!R}^D \times {\rm I\!R}^D \to {\rm I\!R}$ are the mean and covariance functions, respectively,
%
\begin{align*}
 	\mu(\bm{x}) &= \mathbb{E}\!\left[f(\bm{x})\right],\\ 
 	k(\bm{x}, \bm{x}') &= \mathbb{E}\!\left[\left( f(\bm{x}) - \mu(\bm{x}) \right)\left( f(\bm{x}') - \mu(\bm{x}') \right)\right].
\end{align*} 
%
The mean and covariance functions completely characterize the GP prior, and control the a priori behavior of the function $f$. Let $\bm{f}=\left\lbrace f(\bm{x}_n) \right\rbrace_{n=1}^N$, then the resulting prior distribution for $\bm{f}$ is a multivariate Gaussian distribution $\bm{f} \sim \Normal(\bm{\mu}, \bm{K})$, where $\bm{\mu} = \left\lbrace \mu(\bm{x}_n) \right\rbrace_{n=1}^N$ is the mean and $\bm{K}$ the covariance matrix, where $K_{i,j}=k(\bm{x}_i,\bm{x}_j)$. In the following, we focus on zero-mean Gaussian processes, that is set $\mu(\bm{x}) = 0$. The covariance function $k(\bm{x}, \bm{x}')$ might depend on a set of hyperparameters, $\bm{\theta}$, but we will not write this dependency explicitly to ease the notation. The joint distribution of $\bm{f}$ and a new $\tilde{f}$ is also a multivariate Gaussian as,
%
\begin{align*}
p(\bm{f}, \tilde{f})=\Normal \left( \left[ \begin{array}{cc}
\bm{f} \\ 
f^*
\end{array} \right] \,\middle|\, \bm{0},\left[ \begin{array}{cc}
\bm{K}_{\bm{f},\bm{f}} & \bm{k}_{\bm{f},\tilde{f}} \\ 
\bm{k}_{\tilde{f},\bm{f}} & k_{\tilde{f},\tilde{f}}
\end{array} \right] \right),
\end{align*}
%
where $\bm{k}_{\bm{f},\tilde{f}}$ is the covariance between $\bm{f}$ and $\tilde{f}$, and $k_{\tilde{f},\tilde{f}}$ is the prior variance of $\tilde{f}$.

@@

\begin{figure}[!t]
\centering
\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 10mm 0mm, clip]{ch5_fig1_Post_J.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 0mm 5mm 10mm 0mm, clip]{ch5_fig1_Sigma_J.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{ch5_fig1_Cov_J.pdf}}
\vspace{-1mm}
\caption{Mean posterior predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model (dashed red line) and the HSGP model for different number of basis functions $m$, with the boundary factor fixed to a large enough value.}
  \label{fig1_Post_J}
\end{figure}

\begin{figure}
\centering
\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 10mm 0mm, clip]{ch5_fig2_Post_L.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 0mm 5mm 10mm 0mm, clip]{ch5_fig2_Sigma_L.pdf}}
\subfigure{\includegraphics[scale=0.29, trim = 1mm 5mm 3mm 0mm, clip]{ch5_fig2_Cov_L.pdf}}
\vspace{-1mm}
\caption{Mean posterior predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model (dashed red line) and the HSGP model for different values of the boundary factor $c$, with a large enough fixed number of basis functions.}
  \label{fig2_Post_L}
\end{figure}

@@

\begin{figure}
\centering
\begin{tabular}{ c c c }
\arrayrulecolor{darkgray}\hline
c = 1.05 &
\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part1.pdf}
\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part1.pdf} 
\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part1.pdf} & 
\multirow{32}{1cm}{ \includegraphics[scale=0.35, trim = 28mm 30mm 100mm 30mm, clip]{ch5_fig3_legend.pdf}}\\ 
\arrayrulecolor{lightgray}\cline{1-2}
c = 1.1 &
\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part2.pdf} 
\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part2.pdf} 
\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part2.pdf} &\\
\cline{1-2}
c = 1.2 &
\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part3.pdf} 
\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part3.pdf} 
\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part3.pdf} &\\
\cline{1-2}
c = 1.5 &
\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part4.pdf} 
\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part4.pdf} 
\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part4.pdf} & \\
\cline{1-2}
c = 2 &
\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Post_part5.pdf} 
\includegraphics[scale=0.215, trim = 0mm 14mm 0mm 14mm, clip]{ch5_fig3_Sigma_part5.pdf} 
\includegraphics[scale=0.215, trim = 0mm 14mm 5mm 14mm, clip]{ch5_fig3_Cov_part5.pdf} & \\
\cline{1-2}
c = 2.5 &
\includegraphics[scale=0.215, trim = 0mm 4mm 0mm 14mm, clip]{ch5_fig3_Post_part6.pdf}
\includegraphics[scale=0.215, trim = 0mm 4mm 0mm 14mm, clip]{ch5_fig3_Sigma_part6.pdf}
\includegraphics[scale=0.215, trim = 0mm 4mm 5mm 14mm, clip]{ch5_fig3_Cov_part6.pdf} &\\
\arrayrulecolor{darkgray}\hline
\end{tabular}
\caption{Posterior mean predictive functions (left), covariance functions (center) and posterior standard deviations (right) of both the exact GP model and the HSGP model for different number of basis functions $m$ and for different values of the boundary factor $c$.}
  \label{fig3_Post_part1}
\end{figure}

@@

\begin{figure}[H]
\centering
\subfigure{\includegraphics[scale=0.42, trim = 10mm 100mm 5mm 30mm, clip]{ch5_fig7_legend.pdf}}\\
\vspace{-3mm}
\subfigure{\includegraphics[scale=0.31, trim = 1mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part1.pdf}}
\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part2.pdf}}
\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part3.pdf}}\\
\vspace{-3mm}
\subfigure{\includegraphics[scale=0.31, trim = 1mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part4.pdf}}
\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part5.pdf}}
\subfigure{\includegraphics[scale=0.31, trim = 21mm 25.5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part6.pdf}}\\
\vspace{-3mm}
\subfigure{\includegraphics[scale=0.31, trim = 1mm 5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part7.pdf}}
\subfigure{\includegraphics[scale=0.31, trim = 21mm 5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part8.pdf}}
\subfigure{\includegraphics[scale=0.31, trim = 21mm 5mm 10mm 19mm, clip]{ch5_fig7_posterior_varing_lscale_part9.pdf}}\\
\caption{Mean posterior predictions of both exact GP and HSGP models, fitted over various datasets drawn from square exponential GP models with different characteristic lengthscales ($\ell$) and same marginal variance ($\alpha$) as the data-generating functions (\textit{true function}). }
  \label{fig7_posterior_varing_lscale_part1}
\end{figure}

@@

@@

The component $f_1(t)$ represents the long-term trends modeled by a GP with squared exponential covariance function,
%
\begin{equation*}
f_1(t) \sim \GP(0,k_1), \hspace{5mm} k_1(t,t') = \alpha_1 \exp\!\!\left(\!-\frac{1}{2} \frac{(t-t')^2}{\ell_1^2}\right), 
\end{equation*}
%
which means the function values $\bm{f}_1=\{f_1(t)\}_{t=1}^T$ are multivariate Gaussian distributed with covariance matrix $\bm{K}_1$, where $K_{1_{t,s}}=k_1(t,s)$, with $t,s=1,\dots,T$. $\alpha_1$ and $\ell_1$ represent the marginal variance and lengthscale, respectively, of this GP prior component. 
The component $f_2(t)$ represents the yearly smooth seasonal pattern, using a periodic squared exponential covariance function (with period 365.25 to match the average length of the year) in a GP model,
%
\begin{equation*}
f_2(t) \sim \GP(0,k_2), \hspace{5mm} k_2(t,t') = \alpha_2 \exp\!\!\left(\!-\frac{2\,\sin^{\!2}\!(\pi(t-t')/365.25}{\ell_2^2}\right).
\end{equation*}

The component $f_3(t)$ represents the weekly smooth pattern using a periodic squared exponential covariance function (with period 7 of length of the week) in a GP model,
%
\begin{equation*}
f_3(t) \sim \GP(0,k_3), \hspace{5mm} k_3(t,t') = \alpha_3 \exp\!\!\left(\!-\frac{2\,\sin^{\!2}\!(\pi(t-t')/7}{\ell_3^2}\right). 
\end{equation*}

The component $f_4(t)$ represents the special days effects, modeled as a horse-shoe prior model \citep{carvalho2010,piironen2017sparsity}:
%
\begin{equation*}
f_4(t) \sim \Normal(0,\lambda^2_t \tau^2), \hspace{5mm} \lambda^2_t \sim \mathcal{C}^{+}(0,1).
\end{equation*}
%
A horse-shoe prior allows for sparse distributed effects. Its global parameter $\tau$ pulls all the weights (effects) globally towards zero, while the thick half-Cauchy tails for the local scales $\lambda_t$ allow some of the weights to escape the shrinkage. Different levels of sparsity can be accommodated by changing the value of $\tau$: for large $\tau$ all the variables have very diffuse prior distributions with very little shrinkage towards zero, but letting $\tau \rightarrow 0$ will shrink all the weights $f_4(t)$ to zero \citep{piironen2016hyperprior}.

@@
